Partridge: An effective system for the automatic classification of the types of academic papers

### 2013-07-17 doc/bcspaper/article.tex
Made some more minor changes to the article formatting

\vspace{-15pt}
\section{Introduction} \label{sec:1}
\vspace{-11pt}
Since the advent of the `Digital Age', the amount of information available to
researchers has been increasing drastically, relevant material is becoming
progressively more difficult to find manually and the need for an automated
information retrieval tool more apparent. There are already a large number of
information retrieval and recommendation systems for scientific publications.
Many of these systems, such as AGRICOLA\footnote{\url{http://agricola.nal.usda.gov/}},
the Cochrane Library\footnote{\url{http://www.thecochranelibrary.com/}} and Textpresso
\footnote{\url{http://www.textpresso.org/}} index only publications from predefined
journals or topics (for the above examples, Agriculture, Biology and
Bioinformatics respectively).  Unfortunately, these domain specific indexing
systems usually only contain a small subset of papers, excluding potentially
crucial literature because it does not quite fit into the subject domain.

The value of these systems to their users is often restricted by the small
proportion of available literature that they index, forcing researchers to use
multiple, domain specific, search engines for their queries.  In contrast,
there are also a number of interdisciplinary indexing systems and online
journals such as arXiv \footnote{\url{http://arxiv.org}} and PloSOne
\footnote{\url{http://plosone.org}} that try to incorporate wide ranges of
papers from as many disciplines as possible. The traits of these systems often
complement those of their domain-specific counterparts; they provide a
comprehensive collection of literature but insufficient filtering and indexing
capabilities, usually based on title and abstract.

However, the document title is just one of the crucial parts of a scientific
paper's structure.  Liakata \emph{et al.} (2012) describe a system for
automatically processing and classifying sentences in a research paper
according to the core scientific concept (CoreSC) that they
describe\cite{Liakata2012}.  There are 11 CoreSCs, including {\em Hypothesis},
{\em Goal}, {\it Background}, {\em Method}, {\em Result} and {\em Conclusion}.
CoreSC labels can be allocated to all sentences in a scientific paper in order
to identify which scientific concept each sentence encapsulates.  SAPIENTA
\footnote{\url{http://www.sapientaproject.com}} is a publicly available machine learning
application which can automatically annotate all sentences in a scientific
paper with their CoreSC labels. It was trained using a corpus of physical
chemistry and biochemistry research papers whose sentences were manually
annotated using the CoreSC\cite{LIAKATA10.644} scheme.  An intelligent 
information retrieval system can use this data to provide better filtering and
search capabilities for researchers.  Partridge implements such context-aware
keyword search, by allowing researchers to search for papers where a term
appears in sentences with a specific CoreSC label (e.g only in {\em Method}
sentences). This can be used to greatly improve both the precision with which
researchers are able to perform searches for scientific literature and the
accuracy of those searches in terms of relevance to the reader.

@@

% in conclusions add how this can give recommendations for Plos and other
% journals 
A script called {\em plosget.py} was written, in order to download the
papers via the PloS RESTful search API \footnote{\url{http://api.plos.org/solr/faq/}}. The number of papers downloaded per
paper type category was as follows: 200 Essay, 99 Correspondence, 107 Case
Study, 200 Perspective, 74 Viewpoint, 93 Opinion, 312 Review, 200 Research.
These formed a corpus of 1285 %1459-174 after removing synopsis. Should double
check the numbers add up papers.

@@

In order to obtain automatic labeling of sentences in a paper with
CoreSC concepts we first needed to convert the papers to a format that
the the CoreSC classifier, SAPIENTA can analyse.  Currently, SAPIENTA
supports the SciXML and PubMed Central DTDs.  Papers in PDF format are
first converted to XML using PDFX %\cite{PDFX} - no paper known yet
, a free service hosted by the University of Manchester
(\cite{Constantin2013}). Once the paper has been split into
sentences, SAPIENTA is used to annotate each sentence with a relevant
CoreSC label. These labels are assigned using a conditional random
fields model based upon a number of features such as the sentence's
location within the paper and pairs and triplets of words found
consecutively within the sentence \cite{Liakata2012}.  The SAPIENTA
system has been trained on a corpus of 265 chemistry and biochemistry
papers and we have done no domain adaptation for the papers in PLoS.

@@

Random forest learning\cite{Breiman2001} was chosen for the construction of the
paper type classifier. This is because random forests are a fast, accurate and
widely accepted learning technology, and the underlying decision trees can be
inspected to understand some of the reasoning behind the predictions made by
the final paper type classifiers. The feature set consisted of the percentage
composition of each of the 11 respective CoreSC labels assigned to the
sentences in each paper i.e. (the percentage of {\em Background} sentences, the
percentage of {\em Hypothesis} sentences, etc.). The random forest learning was
conducted using the Orange data mining library for Python \cite{Curk2005}.  The
parameters for the random forest learner were as the defaults, except with
min\_subsets=5 and same\_majority\_pruning=true.  We used 10-fold cross
validation to estimate the precision, recall and F-measure.

>>>>

\vspace{-15pt}
\section{Introduction} \label{sec:1}
\vspace{-11pt}

Since the advent of the `Digital Age', the amount of information available to
researchers has been increasing drastically, relevant material is becoming
progressively more difficult to find manually and the need for an automated
information retrieval tool more apparent. There are already a large number of
information retrieval and recommendation systems for scientific publications.
Many of these systems, such as
AGRICOLA\footnote{\url{http://agricola.nal.usda.gov/}}, the Cochrane
Library\footnote{\url{http://www.thecochranelibrary.com/}} and
Textpresso\footnote{\url{http://www.textpresso.org/}} index only publications
from predefined journals or topics (for the above examples, Agriculture,
Biology and Bioinformatics respectively).  Unfortunately, these domain specific
indexing systems usually only contain a small subset of papers, excluding
potentially crucial literature because it does not quite fit into the subject
domain.

The value of these systems to their users is often restricted by the small
proportion of available literature that they index, forcing researchers to use
multiple, domain specific, search engines for their queries.  In contrast,
there are also a number of interdisciplinary indexing systems and online
journals such as arXiv\footnote{\url{http://arxiv.org}} and
PloSOne\footnote{\url{http://plosone.org}} that try to incorporate wide ranges
of papers from as many disciplines as possible. The traits of these systems
often complement those of their domain-specific counterparts; they provide a
comprehensive collection of literature but insufficient filtering and indexing
capabilities, usually based on title and abstract.

However, the document title is just one of the crucial parts of a scientific
paper's structure.  Liakata \emph{et al.} (2012) describe a system for
automatically processing and classifying sentences in a research paper
according to the core scientific concept (CoreSC) that they describe
\cite{Liakata2012}.  There are 11 CoreSCs, including {\em Hypothesis}, {\em
Goal}, {\it Background}, {\em Method}, {\em Result} and {\em Conclusion}.
CoreSC labels can be allocated to all sentences in a scientific paper in order
to identify which scientific concept each sentence encapsulates.
SAPIENTA\footnote{\url{http://www.sapientaproject.com}} is a publicly available
machine learning application which can automatically annotate all sentences in
a scientific paper with their CoreSC labels. It was trained using a corpus of
physical chemistry and biochemistry research papers whose sentences were
manually annotated using the CoreSC \cite{LIAKATA10.644} scheme.  An
intelligent information retrieval system can use this data to provide better
filtering and search capabilities for researchers.  Partridge implements such
context-aware keyword search, by allowing researchers to search for papers
where a term appears in sentences with a specific CoreSC label (e.g only in
{\em Method} sentences). This can be used to greatly improve both the precision
with which researchers are able to perform searches for scientific literature
and the accuracy of those searches in terms of relevance to the reader.

@@

% in conclusions add how this can give recommendations for Plos and other
% journals 
A script called {\em plosget.py} was written, in order to download the
papers via the PloS RESTful search API\footnote{\url{http://api.plos.org/solr/faq/}}. The number of papers downloaded per
paper type category was as follows: 200 Essay, 99 Correspondence, 107 Case
Study, 200 Perspective, 74 Viewpoint, 93 Opinion, 312 Review, 200 Research.
These formed a corpus of 1285 %1459-174 after removing synopsis. Should double
check the numbers add up papers.

@@

In order to obtain automatic labeling of sentences in a paper with CoreSC
concepts we first needed to convert the papers to a format that the the CoreSC
classifier, SAPIENTA can analyse.  Currently, SAPIENTA supports the SciXML and
PubMed Central DTDs.  Papers in PDF format are first converted to XML using
PDFX , a free service hosted by the University of Manchester
\cite{Constantin2013}. Once the paper has been split into sentences, SAPIENTA
is used to annotate each sentence with a relevant CoreSC label. These labels
are assigned using a conditional random fields model based upon a number of
features such as the sentence's location within the paper and pairs and
triplets of words found consecutively within the sentence \cite{Liakata2012}.
The SAPIENTA system has been trained on a corpus of 265 chemistry and
biochemistry papers and we have done no domain adaptation for the papers in
PLoS.

@@

Random forest learning \cite{Breiman2001} was chosen for the construction of the
paper type classifier. This is because random forests are a fast, accurate and
widely accepted learning technology, and the underlying decision trees can be
inspected to understand some of the reasoning behind the predictions made by
the final paper type classifiers. The feature set consisted of the percentage
composition of each of the 11 respective CoreSC labels assigned to the
sentences in each paper i.e. (the percentage of {\em Background} sentences, the
percentage of {\em Hypothesis} sentences, etc.). The random forest learning was
conducted using the Orange data mining library for Python \cite{Curk2005}.  The
parameters for the random forest learner were as the defaults, except with
min\_subsets=5 and same\_majority\_pruning=true.  We used 10-fold cross
validation to estimate the precision, recall and F-measure.
